Methods

### 2021-08-06 text/methods.Rmd
methods model formulas

## 1. Underwater visual census database
\noindent We used a published global database of reef fish abundance and sizes collected along belt transects [@Barneche2019]. This database encompasses 9118 transects across 585 sites (within 98 localities) in the Central Indo-Pacific, Central Pacific, Eastern Pacific, Western Indian, Eastern Atlantic, and Western Atlantic. Sites are defined as small islands or stretches of continuous reefs in larger coastlines and localities encompass sites that belong to the same biogeographic sub-provinces [see @Barneche2019]. The database only includes transects at the outer reef slope and with a hard reef bottom. Transects were carried out at a constant depth, parallel to the reef crest. We selected the species inside families for which we have body stoichiometric data, that were at least 7cm to minimize the bias related to the identification of small individuals, and finally we discarded rare species, for which less than 20 individuals were ever recorded across all transects. The dataset then included 1110 species that belong to 25 families (Acanthuridae, Balistidae, Bothidae, Chaetodontidae, Cirrhitidae, Fistulariidae, Haemulidae, Holocentridae, Kyphosidae, Labridae, Lethrinidae, Lutjanidae, Monacanthidae, Mugilidae, Mullidae, Ostraciidae, Pempheridae, Pomacanthidae, Pomacentridae, Sciaenidae, Scorpaenidae, Serranidae, Siganidae, Tetraodontidae, Zanclidae).
Sea surface temperature (SST) for each site was obtained from daily time‐series data from the National Oceanicand Atmospheric Administration of the USA (NOAA) covering a 5‐year period (°C; 0.25° resolution) (@Reynolds2007; available from https://www.esrl.noaa.gov/psd/data/gridded/data.noaa.oisst.v2.highres.html). Further, for each transect (fish community), we calculated the species richness and estimated the total standing stock biomass of fishes by using Bayesian length-weight relationships available from Fishbase [@Froese2013].

@@

## 5. Imputation of herbivory and piscivory

## 4. Regression models
The standing stock biomass of communities is inevitably positively related to all functions because of the additive nature of the quantification and general metabolic theory. Furthermore, because of the known relationship between temperature and parameters related to growth and respiration, functions are expected to be positively correlated with temperature. To model the effect of biomass and sea surface temperature (sst), while simultaneously investigating the correlations between the the five functions, we performed a multivariate Bayesian mixed effect regression the log-transformed functions for transect-level observations ($i$) including site and location effects:

$$\begin{bmatrix}Y_{ExN,i} \\ Y_{ExP,i} \\ Y_{Prod,i} \\ Y_{Herb,i} \\ Y_{Pisc,i} \end{bmatrix} \sim MVNormal\left(\begin{bmatrix} \mu_{ExN,i} \\ \mu_{ExP,i} \\ \mu_{Prod,i} \\ \mu_{Herb,i} \\ \mu_{Pisc,i}\end{bmatrix}, S \right),$$
$$S = \begin{bmatrix}\sigma_{ExN} & 0 & 0 & 0 & 0\\ 
0 & \sigma_{ExP} & 0 & 0 & 0 \\ 
0 & 0 & \sigma_{Prod,i} & 0 & 0\\ 
0 & 0 & 0 & \sigma_{Herb} & 0\\ 
0 & 0 & 0 & 0 & \sigma_{Pisc} \end{bmatrix} 
R 
\begin{bmatrix}\sigma_{ExN} & 0 & 0 & 0 & 0\\ 
0 & \sigma_{ExP} & 0 & 0 & 0 \\ 
0 & 0 & \sigma_{Prod,i} & 0 & 0\\ 
0 & 0 & 0 & \sigma_{Herb} & 0\\ 
0 & 0 & 0 & 0 & \sigma_{Pisc} \end{bmatrix}$$

$$\mu_{\textrm{ExN,i}} = (\beta0_{ExN} +  \delta_{ExN, loc} +  \delta_{ExN, site}) +{\beta1_{ExN}} {x_{log(biomass),i}} + {\beta2_{ExN}} {x_{sst,i}} $$
$$\beta0_{ExN} \sim  student( 3, -3.9, 2.5)\\
\beta0_{ExP} \sim  student( 3, -6.6, 2.5)\\
\beta0_{Prod} \sim  student( 3, -4.3, 2.5)\\
\beta0_{Herb} \sim  student( 3, -2.4, 2.5)\\
\beta0_{Pisc} \sim  student( 3, -4.6, 2.5)\\
R \sim lkj\_corr(1)
$$

\noindent We used non-centered parametrisation for site and location effects and all standard deviations had the following prior: $\sigma \sim student(3, 0, 2.5)$. To further optimize convergence of the model, we sampled from the cholesky decomposed triangular matrix L such that $R = LL^T$. We used flat priors for the model slopes $\beta1, \beta2$.

The posterior distributions of model parameters were estimated using Markov chain Monte Carlo (MCMC) methods by  using four chains of 2,000 samples, including 1,000 samples as a warm‐up, meaning that a total of 4,000 samples  were used to estimate posterior distributions. We used Bayesian R2 to estimate the amount of explained variation of each model. Posterior predictive checks for are provided in Supporting Informationxxxxx.

\noindent The multivariate  and mixed structure of the above-mentioned model allowed us to extract the correlations between the five functions on the location, site, and  We then assessed the covariation between functions, independent of biomass and sst. To do so, we first extracted the median residuals for each function per transect. In some transects, there were no piscivores or herbivores observed. In those cases, we did not include these transects in the analysis. We then quantified the correlations that exist among the different functions using these median residuals. Finally, for the purpose of visualizing the residual variation of functions per locality on a world map, we ran a supplemental model, similar to the model described above but including random effects both per site and locality. We then extracted and plotted the location effects, which can be interpreted as the average variation per locality.

## 4. Effect of community structure on ecosystem functions 
\noindent To investigate the effect of the community structure while still accounting for the effects of standing biomass and sea surface temperature, Additionally, we quantified the transect-level species richness. 
For each log-transformed function we then fitted a Bayesian mixed-effect model with all 12 above-mentioned variables, after verifying that there are no strong correlations between variables (the highest correlation coefficient was 0.5, and 50% of the variable pair correlations varied between -0.1 and 0.2).

<!-- \begin{equation} -->
$$y_\textrm{j} \sim N(\mu_j, \sigma_j),$$
<!-- 	\label{modcom1} -->
<!-- \end{equation}      -->

<!-- \begin{equation} -->
$\begin{array}{c} \mu_{j} = \beta_0 + \beta_1 x_{log(biomass),j} + \beta_2 x_{sst,j}  + \beta_3 x_{richness,j}  + \beta_4 x_{size_m,j}  + \beta_5 x_{size_{2.5\%},j} +  \beta_6 x_{size_{97.5\%},j} + \\ \beta_7 x_{troph_m,j}  + \beta_8 x_{troph_{2.5\%},j} +  \beta_9 x_{troph_{97.5\%},j} + \beta_{10} x_{imm_m,j}  + \beta_{11} x_{imm_{2.5\%},j} +  \beta_{12} x_{imm_{97.5\%},j} \end{array}$     
<!-- 	\label{modcom2} -->
<!-- \end{equation} -->

To compare effects across functions and assess the relative importance of each variable, we standardized all variables prior to model fitting. We fitted all 5 models by using 4 cores, that each had 2000 iterations with a warm-up of 1000 iterations, and used weakly-informative priors [@Burkner2017].

>>>>

## 1. Underwater visual census database
\noindent We used a published global database of reef fish abundance and sizes collected along belt transects [@Barneche2019]. This database encompasses 9118 transects across 585 sites (within 98 localities) in the Central Indo-Pacific, Central Pacific, Eastern Pacific, Western Indian, Eastern Atlantic, and Western Atlantic. Sites are defined as small islands or stretches of continuous reefs in larger coastlines and localities encompass sites that belong to the same biogeographic sub-provinces [see @Barneche2019]. The database only includes transects at the outer reef slope and with a hard reef bottom. Transects were carried out at a constant depth, parallel to the reef crest. We selected the species inside families for which we have body stoichiometric data, that were at least 7cm to minimize the bias related to the identification of small individuals, and finally we discarded rare species, for which less than 20 individuals were ever recorded across all transects. The dataset then included 1110 species that belong to 25 families (Acanthuridae, Balistidae, Bothidae, Chaetodontidae, Cirrhitidae, Fistulariidae, Haemulidae, Holocentridae, Kyphosidae, Labridae, Lethrinidae, Lutjanidae, Monacanthidae, Mugilidae, Mullidae, Ostraciidae, Pempheridae, Pomacanthidae, Pomacentridae, Sciaenidae, Scorpaenidae, Serranidae, Siganidae, Tetraodontidae, Zanclidae).
Sea surface temperature (SST) for each site was obtained from daily time‐series data from the National Oceanicand Atmospheric Administration of the USA (NOAA) covering a 5‐year period (°C; 0.25° resolution) (@Reynolds2007; available from https://www.esrl.noaa.gov/psd/data/gridded/data.noaa.oisst.v2.highres.html). Further, for each transect (fish community), we calculated the species richness and estimated the total standing stock biomass of fishes by using Bayesian length-weight relationships available from Fishbase [@Froese2013]. All data wrangling and analyses were performed in the software program R (version 4.0.2; R core team 2020).

@@

## 4. Regression models

### 4.1 Imputation of herbivory and piscivory
In the underwater visual transect database, xx transects did not contain herbivores and xx transects did not contain piscivores, yielding false zeros for herbivory and piscivory, respectively. We considered these zeros as missing values because the absence of the observation of a herbivore or piscivore does not prove the actual absence of herbivores and piscivores. In fact, it is highly unlikely that a coral reef fish community contains no herbivores and piscivores at all. To avoid removing all transects with missing values for herbivory or piscivory (n = xx) from our database, we imputed the missing values by fitting two Bayesian regression models using the observed herbivory and piscivory rates and then imputing the missing values by taking the means of posterior predictions. In other words, we used the median predictor values corresponding to missing responses to impute herbivory and piscivory rates. As predictor variables, we used the community structure variables described above and sea surface temperature, all standardized. Additionally, we included random effects of sites and localities on the intercepts.
Both models thus had the following model structure:
$$ y_i \sim normal(\mu_i, \sigma) $$
$$\begin{array}{c}
\mu_{i} = (\beta0 + \delta_{loc} +  \delta_{site}) + \beta_1 x_{log(biomass),i} + \beta_2 x_{sst,i}  + \beta_3 x_{richness,i}  + \beta_4 x_{size_m,i}  + \beta_5 x_{size_{2.5\%},i} +  \beta_6 x_{size_{97.5\%},i} + \\
\beta_7 x_{troph_m,i}  + \beta_8 x_{troph_{2.5\%},i} +  \beta_9 x_{troph_{97.5\%},i} + \beta_{10} x_{imm_m,i}  + \beta_{11} x_{imm_{2.5\%},i} +  \beta_{12} x_{imm_{97.5\%},i} \end{array}$$,

where $i$ represents the transects, $\beta_0$ is the intercept, $\beta_1$ to $\beta_{12}$ are the fixed effect parameters, $\delta_{loc}$ is the random effect of the locality, $\delta_{site}$ is the random effect of the site, biomass represents the total standing stock biomass of the community, sst the sea surface temperature, richness is the species richness, size is the total length, troph is the trophic level, imm is the immaturity, and $m$, $2.5\%$, and $97.5\%$ respectively represent the 50\%, 2.5\%, and 97.5\% quantiles across the fish community.
For these models, we used uninformative priors for the fixed effect parameters ($\beta_{1-12}$) and weakly informative priors for the standard deviations of residuals and random effects (ADD THE PRIOR).
To insure that the performed Imputationdidn't change main results of the downstream regression models, we repeated all below-mentioned models on a subset of the transect data, excluding the original missing values. Results of models are shown in the supplementary information xxxxxxx.

### 4.2 Regressions with biomass and SST, and estimates of correlation

The standing stock biomass of communities is inevitably positively related to all functions because of the additive nature of the quantification. The relationships between functions and biomass are likely to be allometric following metabolic theory. Therefore, we log-transformed each function and biomass in all regression analyses. Furthermore, because of the known relationship between temperature and parameters related to growth and respiration [@Morais2018; refMETABOLIC], functions are hypothesised to be affected by  temperature. To model the effect of biomass and SST, while simultaneously investigating the correlations between the the five functions (independent of biomass and SST), we performed a multivariate Bayesian mixed effect regression model:

$$\begin{bmatrix}log(Y_{ExN,i}) \\ log(Y_{ExP,i}) \\ log(Y_{Prod,i}) \\ log(Y_{Herb,i}) \\ log(Y_{Pisc,i}) \end{bmatrix} \sim MVNormal\left(\begin{bmatrix} \mu_{ExN,i} \\ \mu_{ExP,i} \\ \mu_{Prod,i} \\ \mu_{Herb,i} \\ \mu_{Pisc,i}\end{bmatrix}, S \right),$$
$$S = \begin{bmatrix}\sigma_{ExN} & 0 & 0 & 0 & 0\\
0 & \sigma_{ExP} & 0 & 0 & 0 \\
0 & 0 & \sigma_{Prod,i} & 0 & 0\\
0 & 0 & 0 & \sigma_{Herb} & 0\\
0 & 0 & 0 & 0 & \sigma_{Pisc} \end{bmatrix}
R
\begin{bmatrix}\sigma_{ExN} & 0 & 0 & 0 & 0\\
0 & \sigma_{ExP} & 0 & 0 & 0 \\with the exclusion of
0 & 0 & \sigma_{Prod,i} & 0 & 0\\
0 & 0 & 0 & \sigma_{Herb} & 0\\
0 & 0 & 0 & 0 & \sigma_{Pisc} \end{bmatrix}$$

$$\mu_{\textrm{ExN,i}} = (\beta0_{ExN} +  \delta_{ExN, loc} +  \delta_{ExN, site}) +{\beta1_{ExN}} {x_{log(biomass),i}} + {\beta2_{ExN}} {x_{sst,i}} $$
$$\mu_{\textrm{ExP,i}} = (\beta0_{ExP} +  \delta_{ExP, loc} +  \delta_{ExP, site}) +{\beta1_{ExP}} {x_{log(biomass),i}} + {\beta2_{ExP}} {x_{sst,i}} $$
$$\mu_{\textrm{Prod,i}} = (\beta0_{Prod} +  \delta_{Prod, loc} +  \delta_{Prod, site}) +{\beta1_{Prod}} {x_{log(biomass),i}} + {\beta2_{Prod}} {x_{sst,i}} $$
$$\mu_{\textrm{Herb,i}} = (\beta0_{Herb} +  \delta_{Herb, loc} +  \delta_{Herb, site}) +{\beta1_{Herb}} {x_{log(biomass),i}} + {\beta2_{Herb}} {x_{sst,i}} $$
$$\mu_{\textrm{Pisc,i}} = (\beta0_{Pisc} +  \delta_{Pisc, loc} +  \delta_{Pisc, site}) +{\beta1_{Pisc}} {x_{log(biomass),i}} + {\beta2_{Pisc}} {x_{sst,i}} $$,

where i is the transect, Y_{ExN,i} is N excretion, Y_{ExP,i} is P excretion, Y_{Prod,i} is biomass production, Y_{Herb,i} is herbivory, Y_{ExN,i} is piscivory, \sigma represents the residual error of each function (ExN, ExP, Prod, Herb, and Pisc), R is the correlation matrix of the residuals. Locality- and site-level effects are also structured including covariation among functions. There are thus three correlation matrices in total, or in other words the model will estimate the correlation between functions (independent of biomass and SST) on three levels: locality-level, site-level, and transect-level.

\noindent We used non-centered parametrisation for site and location effects and all standard deviations had the following prior: $\sigma \sim student(3, 0, 2.5)$. We used a lkj_corr prior for each of the three correlation matrices ($R \sim lkj\_corr(1)$) (ADD REF). We used weakly-informative normal priors for the model slopes ($\beta1 \sim normal(1,1)$, $\beta2 \sim normal(0,1)$).

### 4.3 Intercepts-only models
Additionally, we performed a multivariate intercept-only model to estimate the functions per locality, and a univariate intercept-only model to estimate the multifunction per locality. The model structures were similar as the above-mentioned models, the only difference being the exclusion of the predictor variables (sst and biomass).
Further, we fitted a similar univariate model to predict the multifunction per locality:
$$\mu_{\textrm{multi,i}} = \beta0_{multi} +  \delta_{multi, loc} +  \delta_{multi, site} $$

### 4.4 Effect of community structure on ecosystem functions
\noindent To investigate the effect of the community structure while still accounting for the effects of standing biomass and sea surface temperature, Additionally, we quantified the transect-level species richness.
For each log-transformed function we then fitted a Bayesian mixed-effect model with all 12 above-mentioned variables, after verifying that there are no strong correlations between variables (the highest correlation coefficient was 0.5, and 50% of the variable pair correlations varied between -0.1 and 0.2).

$$y_\textrm{j} \sim N(\mu_j, \sigma_j),$$

$\begin{array}{c} \mu_{j} = \beta_0 + \beta_1 x_{log(biomass),j} + \beta_2 x_{sst,j}  + \beta_3 x_{richness,j}  + \beta_4 x_{size_m,j}  + \beta_5 x_{size_{2.5\%},j} +  \beta_6 x_{size_{97.5\%},j} + \\ \beta_7 x_{troph_m,j}  + \beta_8 x_{troph_{2.5\%},j} +  \beta_9 x_{troph_{97.5\%},j} + \beta_{10} x_{imm_m,j}  + \beta_{11} x_{imm_{2.5\%},j} +  \beta_{12} x_{imm_{97.5\%},j} \end{array}$

### 4.5 Model specifications
We ran all Bayesian reggression models using the R package *brms* which uses the stan algorithm.
The posterior distributions of model parameters were estimated using Markov chain Monte Carlo (MCMC) methods by  using four chains of 2,000 samples, including 1,000 samples as a warm‐up, meaning that a total of 4,000 samples were used to estimate posterior distributions.
For all models we inspected the MCMC chains and the Rhat for convergence and peformed posterior predictive checks (Appendix xxxx).
We used Bayesian R2 to estimate the amount of explained variation of each model.
